---
title: "Ex2 for the students"
author: "Lefkios Paikousis"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 2:
Objective: Perform data filtering and summarisation to extract specific information.

## Load the necessary packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(gapminder)
```


## View the LAST 6/ 10/ 20 rows of the `gapminder`dataset.
HINT: Use the function `tail()`

```{r}
tail(gapminder, 20)
```

## Calculate the `minimum`, `maximum`, and `median` values for the 'lifeExp' variable
HINT: You may use the `summary()` function to get all the summary statistics
```{r}
min(gapminder$lifeExp)
median(gapminder$lifeExp)
max(gapminder$lifeExp)

summary(gapminder$lifeExp)
```


## Calculate the average GDP per capita for the Greece across all years

```{r}

gapminder |> 
  filter(country == 'Greece') |> 
  summarise(mean_gdp = mean(gdpPercap))

```

# Find the 5 countries with the lowest Life expectancy in the year 2002.

```{r}
gapminder |> 
  filter(year == 2002) |> 
  arrange(lifeExp) |> 
  head(n = 5)
```

# Calculate the total population for Europe in every year

```{r}
gapminder |>
  filter(continent == "Europe") |> 
  group_by(continent, year) |> 
  summarise(
    total_population = sum(pop)
  )

```


