---
title: "Ex2-Filters Logical Operators"
author: "Lefkios Paikousis"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise:
Objective: Perform more advanced data filtering to extract specific information.


## Load the necessary packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(gapminder)
```


## Find the countries that have life expectancy greater or equal to 80 years in 2007

```{r}

gapminder |> 
  filter(year == 2007, lifeExp >=80) 
```

# How many in each continent?

```{r}

gapminder |> 
  filter(year == 2007, lifeExp >=80) |> 
  count(continent)
```


# Countries in Europw, where the gdp per capita in 2007 is lower than that of Europe's average gdp per capita in 2007
HINT: first find the mean gdp per capita of Europe in 2007, and use that to filter the dataset    


```{r}
gapminder |> 
  filter(year == 2007, continent == "Europe") |>
  group_by(continent) |> 
  summarise(avg_gdp = mean(gdpPercap))

gapminder |> 
  filter(year == 2007, continent == "Europe", gdpPercap < 25054)

```


# Exlude the countries `Japan`, `Canada`, `Greece` from the dataset
HINT: Search google with the term `dplyr not in` ans see the first couple of results     


```{r}

my_countries <- c("Japan", "Canada", "Greece")

gapminder |> 
  filter( !country %in% my_countries)

```


## Find the country with the highest life expectancy in the year 2007.

HINT: You will use the `max()` function inside the `filter()`
```{r}
gapminder |> 
  filter( year == 2007, lifeExp == max(lifeExp) )

```


