---
title: "Ex3"
author: "Lefkios Paikousis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Exercise 3:
Objective: Visualize the data using ggplot2 to gain insights.


## Load the necessary packages

```{r}
library(tidyverse)
library(gapminder)
```


## Create a scatter plot using ggplot2 to show the relationship between 'gdpPercap' and 'lifeExp' variables for the year 2007. Add color to the scatter plot points based on the continent.

```{r}
gapminder |> 
  filter(year == 2007) |> 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent))

```

## Create a line plot using ggplot2 to show the trend of average life expectancy across the years for a specific country.

```{r}
gapminder |> 
  filter(country == "Ireland") |> 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line(aes(color = country))

```

## Create a bar plot to compare the total population of different continents in the year 2007.

```{r}
gapminder |> 
  filter(year == 2007) |> 
  group_by(continent) |> 
  summarise(total_population = sum(pop)) |> 
  ggplot(aes(x = continent, y = total_population))+
  geom_col(width = 0.5)+
  coord_flip()+
  scale_y_continuous(labels = scales::label_number(suffix = "B", scale = 1e-9))

```


## Create a boxplot to compare the distribution of life expectancy across different continents in the year 2007.

```{r}
gapminder |> 
  filter(year == 2007) |> 
  ggplot(aes(x = continent, y = lifeExp))+
  geom_boxplot()+
  geom_point(position = position_jitter(width = 0.15))

```




