---
title: "Exercise 4"
author: "Lefkios Paikousis"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 4:
Objective: Perform data manipulation using dplyr to answer specific questions.

## Load the necessary packages

```{r}
library(tidyverse)
library(gapminder)
```


## For each continent, for the year 2007 calculate:

- the average life expectancy 
- the average gdp Per Capita, 
- total population in millions
- number of countries in each continent


```{r}
gapminder |> 
  filter(year == 2007) |> 
  group_by(continent) |> 
  summarise(
    countries = n(),
    countries2 = n_distinct(country),
    avg_gdpPercap = mean(gdpPercap),
    avg_lifeExp = mean(lifeExp),
    total_population_millions = sum(pop) / 1e6
  )

```

## Create 2 new variables
 - 'population_millions' that represents the population in millions.
 - the actual `GDP` in the country. This is the `gdpPercap` multiplied with `population`

```{r}
gapminder |> 
  mutate(
    pop_millions = pop / 1000000,
    GDP = gdpPercap * pop
  )
```

## Get the average GDP in billions for the `Americas` as a continent, in year 2002
- Then compare this, to what would have been, if you excluded the country with the highest GDP
- Which country in the Americas has the highest GDP

```{r}
gapminder |> 
  filter(year == 2002, continent == "Americas") |> 
  mutate(
    gdp = gdpPercap * pop
  ) |> 
  filter(gdp != max(gdp)) |> 
  summarise(
    avg_gdp = mean(gdp) / 1000000000
  ) 
```







