---
title: "Penguin Plot"
author: "Lefkios Paikousis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plotting the association of `flipper length` and body mass`

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(palmerpenguins)
```

- Check the dataset out
```{r}
penguins
```

- See the variables and their type

```{r}
glimpse(penguins)

```

- View trhe full dataset in the Viewer
```{r}
View(penguins)

```

- Information on the dataset
```{r}
?penguins
```


- Step 1. The blank canvas
```{r}
ggplot(data = penguins)
```

- map visual properties (aesthetics) to axis

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
)

```

- geom: the geometrical object that a plot uses to represent data
```{r}


ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point()

```


- Adding aesthetics and layers
```{r}

ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) +
  geom_point()
```


-Add a line representing the association

```{r}


ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) +
  geom_point() +
  geom_smooth()
```

- We have 3 lines, one for each species
The aesthetic mappings are defined in ggplot(), at the global level,they’re passed down to each of the subsequent geom layers of the plot

- Move the `aes(color = species)` in the `geom_smooth()`
```{r}

ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(aes(color = species)) +
  geom_smooth(method = "lm")
```


- Add Titles on axis and on the chart
```{r}

ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  )
```


# Other type of plots

- Bar plot. Height of the bar is the count of the observations

```{r}


ggplot(penguins, aes(x = fct_infreq(species))) +
  geom_bar(width = 0.5)

```

- Histogram

```{r}

ggplot(penguins, aes(x = body_mass_g)) +

  geom_histogram(binwidth = 200)
```

- Bar chart by Species
```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar()
```

- Fill the bars with proportions
```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(position = "fill")
```


# Facets

- By island

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = species)) +
  facet_wrap(~island)
```


