---
title: "Reshape (pivoting the data) using {tidyr}"
author: "Lefkios Paikousis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)

```

# Read in a .csv file of life expectancy
- Transform it in a long format

```{r}
dta <- readr::read_csv("data/life_expectancy2.csv")

dta_long <- dta |> 
  pivot_longer(-c(country, continent),
               names_to = "year",
               values_to = "lifeExp"
  )

glimpse(dta_long)

```

# Get the last 5 years of data and take it back to wide format, with years as the columns
```{r}
dta_long |> 
  filter(year < "2005") |> 
  pivot_wider(
    names_from = year,
    values_from = lifeExp
  )
```


# The mean gdp per capita, by Year `AND` Continent, in a wide format

```{r}
dta_long |> 
  filter(year < "2005") |> 
  group_by(year, continent) |> 
  summarise(lifeExp = mean(lifeExp)) |> 
  pivot_wider(names_from = year, values_from = lifeExp) 

```


# Use it with simple counts to check the quality of the dataset

- Question: DO we have the same number of countries in each continent, in each year?
- Check only the years before 2005 (to make it easier)

```{r}

dta_long |> 
  filter(year < "2005") |> 
  group_by(year, continent) |> 
  summarise(
    n_countries = n_distinct(country)
  ) |> 
  pivot_wider(names_from = year, values_from = n_countries)
  
```



# Add the europe sub region in the european countries
```{r}
europe_countries <- readxl::read_xlsx("data/europe.xlsx")

glimpse(europe_countries)
```


```{r}
dta_long |> 
  filter(year == 2020) |> 
  left_join(
    europe_countries, 
    by = c("country" = "Country") ) |> 
  filter(!is.na(Subregion)) |> 
  rename(pop_2020 = `Population(2020)`) 

```



